Sampling requirements to detect ecosystem change

### 2020-03-05 R/manuscript_draft.Rmd
edit figures and remove cases with several n2 init for the case study

p1  <- ggplot(myoutput_summary, aes(x=number_samples,y=mean_distance)) + geom_line()  + geom_ribbon(aes(ymin = mean_distance - 1.96*sd_distance, ymax = mean_distance + 1.96*sd_distance), fill = "grey80") +geom_line(size=1) + coord_cartesian(ylim = c(2.2, 50))  + 
  facet_wrap(~approach,
             labeller = label_glue('({.l}) {approach} sampling ')) + 
  theme_bw() +
  labs(x="Total number of samples analyzed", y="Distance to real changepoint") #+
 # theme(legend.position = c(0.93, 0.72),
#       legend.background = element_rect(fill = "#ffffffaa", colour = NA),
#        legend.title=element_text(size=7), 
 #       legend.text=element_text(size=8)) +
#  guides(pch=guide_legend(title="Initial number\nof sample"))

@@

We found that random sampling performed the worst, as changepoint analysis was left to chance. Regular sampling provided good estimates from 8 samples (0-3 sample of absolute error), but detecting the true changepoint depended on the interval falling close to the true changepoint. Iterative sampling, where 5 samples are initially sampled, then other samples are slowly added, performed the best, as no more than 9 samples were ever necessary to get the true changepoint (Fig. \ref{fig:case_study_number_sample} (c)). We show how iterative sampling changes the ordination, and the resulting changepoint, as more samples are added (Fig. \ref{fig:case_study_iterative_sampling}) <!--Note that when the initial number of samples was set to 7 Fig. \ref{fig:case_study_number_sample} (c), square markers), the distance to the real changepoint was initially very high. Without any prior, the initial number sampled fell far away from the true changepoint.-->

@@

```{r case study read data output, include=FALSE}
myoutput <- read.table(paste0(getwd(),"/Output/output_changepoint_varese.txt"))
myoutput$method <- factor(myoutput$method, levels=c('Random', 'Regular', 'Iterative'))

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width = 7.2, fig.height= 3,fig.cap='WHY ARE THERE NEGATIVE VALUES HERE? Distance to real changepoint per total number of samples analyzed, following (a) random sampling, (b) regular sampling, and (c) iterative sampling. Total number of samples was set between 5 and 30, out of the 74 initial time series.  \\label{fig:case_study_number_sample}'}
p1 <- ggplot(myoutput[myoutput$target_cpt==1,], aes(final_n,diff_real, pch=factor(n2_init))) + 
  geom_hline(aes(yintercept=0), col=adjustcolor("black", alpha.f = .4)) + 
  geom_point(alpha = .3) +
  facet_wrap(~method,
             labeller = label_glue('({.l}) {method} sampling ')) + 
  theme_bw() +
  labs(x="Total number of samples analyzed", y="Distance to real changepoint") +
  theme(legend.position = c(0.93, 0.72),
        legend.background = element_rect(fill = "#ffffffaa", colour = NA),
        legend.title=element_text(size=7), 
        legend.text=element_text(size=8)) +
  guides(pch=guide_legend(title="Initial number\nof sample"))

>>>>

p1  <- ggplot(myoutput_summary, aes(x=number_samples,y=mean_distance)) + geom_line()  + geom_ribbon(aes(ymin = mean_distance - 1.96*sd_distance, ymax = mean_distance + 1.96*sd_distance), fill = "grey80") +geom_line(size=1) + coord_cartesian(ylim = c(2.2, 50))  + 
  facet_wrap(~approach,
             labeller = label_glue('({.l}) {approach} sampling ')) + 
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  labs(x="Total number of samples analyzed", y="Distance to real changepoint") #+
 # theme(legend.position = c(0.93, 0.72),
#       legend.background = element_rect(fill = "#ffffffaa", colour = NA),
#        legend.title=element_text(size=7), 
 #       legend.text=element_text(size=8)) +
#  guides(pch=guide_legend(title="Initial number\nof sample"))

@@

We found that random sampling performed the worst, as changepoint analysis was left to chance. Regular sampling provided good estimates from 8 samples (0-3 sample of absolute error), but detecting the true changepoint depended on the interval falling close to the true changepoint (i.e., left to chance). Iterative sampling, where 5 samples are initially sampled, then other samples are slowly added, performed the best, as no more than 9 samples were ever necessary to get the true changepoint (Fig. \ref{fig:case_study_number_sample} (c)). We show how iterative sampling changes the ordination, and the resulting changepoint, as more samples are added (Fig. \ref{fig:case_study_iterative_sampling}) <!--Note that when the initial number of samples was set to 7 Fig. \ref{fig:case_study_number_sample} (c), square markers), the distance to the real changepoint was initially very high. Without any prior, the initial number sampled fell far away from the true changepoint.-->

@@

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width = 7.2, fig.height= 3,fig.cap='WHY ARE THERE NEGATIVE VALUES HERE? Distance to real changepoint per total number of samples analyzed, following (a) random sampling, (b) regular sampling, and (c) iterative sampling. Total number of samples was set between 5 and 30, out of the 74 initial time series.  \\label{fig:case_study_number_sample}'}
# Read data
myoutput <- read.table(paste0(getwd(),"/Output/output_changepoint_varese.txt"))
myoutput$method <- factor(myoutput$method, levels=c('Random', 'Regular', 'Iterative'))
myoutput$diff_real <- abs(myoutput$diff_real)

# Change to TRUE if we decide it is worth it to show several initial number of n2. It looks more confusing -- but it would bring more information
several_n2_init = F

# Plot
p1 <- if(several_n2_init) {
  ggplot(myoutput[myoutput$target_cpt==1,], aes(final_n,diff_real, pch=factor(n2_init))) + 
  geom_hline(aes(yintercept=0), col=adjustcolor("black", alpha.f = .4)) + 
  geom_point(alpha = .3) +
  facet_wrap(~method,
             labeller = label_glue('({.l}) {method} sampling ')) + 
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  labs(x="Total number of samples analyzed", y="Distance to real changepoint") +
  theme(legend.position = c(0.93, 0.72),
        legend.background = element_rect(fill = "#ffffffaa", colour = NA),
        legend.title=element_text(size=7), 
        legend.text=element_text(size=8)) +
  guides(pch=guide_legend(title="Initial number\nof sample"))
} else {
  ggplot(myoutput[myoutput$target_cpt==1 & myoutput$n2_init == 5,], aes(final_n,diff_real)) + 
  geom_hline(aes(yintercept=0), col=adjustcolor("black", alpha.f = .4)) + 
  geom_point(alpha = .3) +
  facet_wrap(~method,
             labeller = label_glue('({.l}) {method} sampling ')) + 
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  labs(x="Total number of samples analyzed", y="Distance to real changepoint") 
}
